Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
update

@@

@@

\textbf{Sender Email:} We found that spammers were

\textbf{Sender Age:} Spam user age

@@

\textbf{Sender Birthday:}

@@

\begin{table}
\begin{tabular}{|l|l|r|r|r|}
    \hline
    Feature & Class & Mean & Median & Std. Dev.\\
    \hline
    \multirow{2}{*}{Unique IPs} & Spam & 11.82 & 2 & 52.52\\
    & Ham & 175.51 & 64 & 322.19\\
    \hline
    \multirow{2}{*}{Sender Age} & Spam & 33.52 & 31 & 11.09\\
    & Ham & 24.86 & 22 & 10.30\\
    \hline
    \multirow{2}{*}{Photos} & Spam & 2.09 & 1 & 8.04\\
    & Ham & 15.69 & 6 & 43.40\\
    \hline
    \multirow{2}{*}{Photo Albums} & Spam & 0.91 & 1 & 0.56\\
    & Ham & 1.89 & 1 & 2.43\\
    \hline
    \multirow{2}{*}{Friends} & Spam & 11.82 & 0 & 52.52\\
    & Ham & 23.26 & 10 & 58.91\\
    \hline
    \multirow{2}{*}{Recipient Age} & Spam & 39.98 & 41 & 12.69\\
    & Ham & 25.02 & 22 & 10.70\\
    \hline
    \multirow{2}{*}{Birth Day} & Spam & 12.31 & 10 & 8.52\\
    & Ham & 15.01 & 15 & 8.92\\
    \hline
    \multirow{2}{*}{Birth Month} & Spam & 5.85 & 5 & 3.33\\
    & Ham & 6.38 & 6 & 3.48\\
    \hline
    \multirow{2}{*}{Description Length} & Spam & 267.30 & 87 & 483.19\\
    & Ham & 439.33 & 267 & 613.26\\
    \hline
    \multirow{2}{*}{\% Fields Completed} & Spam & 22.15 & 14.00 & 30.75\\
    & Ham & 79.29 & 86.00 & 32.85\\
    \hline
\end{tabular}
\caption{Feature statistics}
\label{featsat}
\end{table}

@@

\subsection{Features for Classification}
In choosinnng

@@

%feature extraction tools: perl, spark, lucene analyzers
%- feature extraction pipeline is parallelized / distributed using Spark
%- it produces a feature matrix used later for training classifiers
%- Stage 1 generates a dictionary sorted by document frequencies in descending order
%- Stage 2 create bag-of-words representation of the examples, as well as “expert-features” along with the documents.
%
%- two feature sets: one just with bag-of-words. the other with expert features chosen by a domain expert (site owner)
%
%classification:
%- implemented 3 models: naive bayes, linear regression, and SVM.
%- used scalanlp for SVM: Pegasos: Primal Estimated sub-GrAdient SOlver for SVM Extended with Wang, Crammer, Vucetic's work for Multiclass. The optimizer runs a stochastic subgradient descent on the primal objective using batches provided.
%- used scalanlp for naive bayes
%- used matlab for linear regression with ridge regularization.
%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
%
%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
%
%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
%
%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
%

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:day}
figure \ref{fig:month}
figure \ref{fig:drop}
figure \ref{fig:email}
figure \ref{fig:friends}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:roc-25}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/dob-day.pdf}
    \caption{Date of month from birthday submitted by user at signup.}
    \label{fig:day}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/email.pdf}
    \caption{Domain name of email provided at signup. Users cannot send messages before 
        activating their account via a link sent to their email address.}
    \label{fig:email}
\end{figure}

>>>>

\setcounter{page}{1}
\pagenumbering{arabic}

@@

We chose message and user account features to analyze primarily by the amount of data that we had for each them, while making sure to include the most vital account information. Table \ref{tab:feats} provides averages on a number of the account statistics that we examined. 
\begin{table}
\begin{tabular}{|l|l|r|r|r|}
    \hline
    \textbf{Feature} & \textbf{Label} & \textbf{Mean} & \textbf{Median} & \textbf{Std. Dev.}\\
    \hline
    \multirow{2}{*}{Unique IPs} & Spam & 11.82 & 2 & 52.52\\
    & Ham & 175.51 & 64 & 322.19\\
    \hline
    \multirow{2}{*}{Sender Age} & Spam & 33.52 & 31 & 11.09\\
    & Ham & 24.86 & 22 & 10.30\\
    \hline
    \multirow{2}{*}{Photos} & Spam & 2.09 & 1 & 8.04\\
    & Ham & 15.69 & 6 & 43.40\\
    \hline
    \multirow{2}{*}{Photo Albums} & Spam & 0.91 & 1 & 0.56\\
    & Ham & 1.89 & 1 & 2.43\\
    \hline
    \multirow{2}{*}{Friends} & Spam & 11.82 & 0 & 52.52\\
    & Ham & 23.26 & 10 & 58.91\\
    \hline
    \multirow{2}{*}{Recipient Age} & Spam & 39.98 & 41 & 12.69\\
    & Ham & 25.02 & 22 & 10.70\\
    \hline
    \multirow{2}{*}{Birth Day} & Spam & 12.31 & 10 & 8.52\\
    & Ham & 15.01 & 15 & 8.92\\
    \hline
    \multirow{2}{*}{Birth Month} & Spam & 5.85 & 5 & 3.33\\
    & Ham & 6.38 & 6 & 3.48\\
    \hline
    \multirow{2}{*}{Profile Desc. Len.} & Spam & 267.30 & 87 & 483.19\\
    & Ham & 439.33 & 267 & 613.26\\
    \hline
    \multirow{2}{*}{\% Fields Completed} & Spam & 22.15 & 14.00 & 30.75\\
    & Ham & 79.29 & 86.00 & 32.85\\
    \hline
\end{tabular}
\caption{Feature statistics}
\label{tab:feats}
\end{table}

@@

\textbf{Sender Email:} Users must provide email addresses upon signup. The site does not allow users to contact other 
members before verifying that the email address provided is valid, via an account activation link sent to the signup 
address. Subsequent changes to a user's email address on file require a similar confirmation process. Thus, we can be certain 
that the email accounts that we extracted were in use by the senders, at least at the time of registration.

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/email.pdf}
    \caption{Domain name of email provided at signup.}
    \label{fig:email}
\end{figure}

In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no explanation for 
this, figure \cite{fig:ysmap} shows that Ghana and Nigeria alone account for 38\% of the IP addresses associated with these 
Yahoo accounts. This popularity of Yahoo among West African users has been noted before\cite{burrell}, and appears to be 
due to the early penetration of Yahoo services in the region.

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/yahoo-spam.pdf}
    \caption{IP distribution for scam accounts with Yahoo email addresses.}
    \label{fig:yspam}
\end{figure}

\textbf{Sender Age:} Spam user age tended to be higher

\textbf{Sender Birthday:} We analyzed

figure \ref{fig:day}
figure \ref{fig:month}
figure \ref{fig:drop}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/dob-day.pdf}
    \caption{Date of month from birthday submitted by user at signup.}
    \label{fig:day}
\end{figure}

@@

@@

@@

@@

In order to train our classifiers, we had to create a feature matrix from the message and account data. We did this using 
Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document frequencies in descending 
order, mapping each feature (word or expert feature value) to an integer key. The second step was to create a sparse 
matrix representation of each document (in this case, spam or ham message) and the associated features. Spark provides a 
parallelized in-memory computation framework that dramatically reduced the time necessary to extract these features, 
build the dictionary and generate the feature matrix.

We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
message text, we used the following features:
\begin{itemize}
\item Sex of sender
\item Age of sender
\item Age of recipient
\item Account lifetime
\item Month of birth 
\item Whether sender has friends (boolean)
\item Whether sender has photo(s) (boolean)
\item Profile country and IP-detected country match (boolean)
\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) (boolean)
\end{itemize}

%classification:
%- implemented 3 models: naive bayes, linear regression, and SVM.
%- used scalanlp for SVM: Pegasos: Primal Estimated sub-GrAdient SOlver for SVM Extended with Wang, Crammer, Vucetic's work for Multiclass. The optimizer runs a stochastic subgradient descent on the primal objective using batches provided.
%- used scalanlp for naive bayes
%- used matlab for linear regression with ridge regularization.
%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
%
%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
%
%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
%
%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
%

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:drop}
figure \ref{fig:friends}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:roc-25}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

@@
